---
title: 'Text mining, sentiment analysis, and visualization'
author: "William Mailil"
date: "05-08-2026"
output: "html_document"
---
## Loading libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(here)
library(pdftools)
library(tidytext)
library(textdata) 
library(ggwordcloud)
```

## Getting the got.pdf:
```{r get-document}
got_path <- here("data","got.pdf")
got_text <- pdf_text(got_path)
```

## Testing by extracting a page
```{r single-page}
#I can read the pdf, but it also reads white-space like \n\ which I want to filter
got_p9 <- got_text[9]
got_p9
```
## Splitting up the book
```{r split-lines}
got_df <- data.frame(got_text) %>% 
  mutate(text_full = str_split(got_text, pattern = '\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 
```

## Getting the tokens in tidy format


```{r tokenize}
got_tokens <- got_df %>% 
  unnest_tokens(word, text_full)

got_tokens

```

```{r count-words}
got_wc <- got_tokens %>% 
  count(word) %>% 
  arrange(-n)
got_wc
```

## Removing stop words:
```{r stopwords}
got_stop <- got_tokens %>% 
  anti_join(stop_words) %>% 
  select(-got_text)

#Counting again 
got_swc <- got_stop %>% 
  count(word) %>% 
  arrange(-n)

got_swc
```

```{r skip-numbers}
got_no_numeric <- got_stop %>% 
  filter(is.na(as.numeric(word)))

got_no_numeric

# I think this step is unnecessary with this text and the result is worse/wrong

```

## A word cloud of IPCC report words (non-numeric)
```{r wordcloud-prep}
# I want to compare got_no_numeric and got_swc

length(unique(got_no_numeric$word))

length(unique(got_swc$word))

# Ok I was wrong got_no_numeric seems to work as intended 


got_top100 <- got_no_numeric %>% 
  count(word) %>% 
  arrange(-n) %>% 
  head(100)
got_top100

# no mention of "grease", very disappointing
```

```{r wordcloud}
got_cloud <- ggplot(data = got_top100, aes(label = word)) +
  geom_text_wordcloud() +
  theme_minimal()

got_cloud
```

```{r wordcloud-pro}
ggplot(data = got_top100, aes(label = word, size = n)) +
  geom_text_wordcloud_area(aes(color = n), shape = "star") +
  scale_size_area(max_size = 12) +
  scale_color_gradientn(colors = c("darkgreen","blue","red")) +
  theme_minimal()
```


## Sentiment analysis
```{r afinn}
get_sentiments(lexicon = "afinn")

afinn_pos <- get_sentiments("afinn") %>% 
  filter(value %in% c(3,4,5))

afinn_pos
```

```{r bing}
get_sentiments(lexicon = "bing")

```


```{r nrc}
get_sentiments(lexicon = "nrc")

```

## Sentiment analysis with afinn: 
```{r bind-afinn}
got_afinn <- got_stop %>% 
  inner_join(get_sentiments("afinn"))
got_afinn
```

```{r count-afinn}
got_afinn_hist <- got_afinn %>% 
  count(value)

ggplot(data = got_afinn_hist, aes(x = value, y = n)) +
  geom_col(aes(fill = value)) +
  theme_bw()
# I am not surprised
```

```{r afinn-2}
got_afinn2 <- got_afinn %>% 
  filter(value == 2)
```

```{r afinn-2-more}
unique(got_afinn2$word)

got_afinn2_n <- got_afinn2 %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = fct_reorder(factor(word), n))

ggplot(data = got_afinn2_n, aes(x = word, y = n)) +
  geom_col() +
  coord_flip() +
  theme_bw()

# I can't really see anything, it's too clustered
```

```{r summarize-afinn}
got_summary <- got_afinn %>% 
  summarize(
    mean_score = mean(value),
    median_score = median(value)
  )
got_summary
# Makes sense
```


## NRC lexicon for sentiment analysis
```{r bind-bing}
got_nrc <- got_stop %>% 
  inner_join(get_sentiments("nrc"))
# Doesn't work
```


```{r check-exclusions}
got_exclude <- got_stop %>% 
  anti_join(get_sentiments("nrc"))

got_exclude_n <- got_exclude %>% 
  count(word, sort = TRUE)

head(got_exclude_n)
```


```{r count-bing}
got_nrc_n <- got_nrc %>% 
  count(sentiment, sort = TRUE)

ggplot(data = got_nrc_n, aes(x = sentiment, y = n)) +
  geom_col(aes(fill = sentiment))+
  theme_bw()
```

```{r count-nrc}
got_nrc_n5 <- got_nrc %>% 
  count(word,sentiment, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  top_n(5) %>% 
  ungroup()

got_nrc_gg <- ggplot(data = got_nrc_n5, aes(x = reorder(word,n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, ncol = 2, scales = "free") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Word", y = "count")

got_nrc_gg

# Saving
ggsave(plot = got_nrc_gg, 
       here("figures","got_nrc_sentiment.png"), 
       height = 8, 
       width = 5)

```

```{r nrc-confidence}
# I did this with "lord" instead
conf <- get_sentiments(lexicon = "nrc") %>% 
  filter(word == "lord")

conf
```

## Credits: 
This tutorial is inspired by Allison Horst's Advanced Statistics and Data Analysis.
